---
title: "Induced demand Rebound effect from ebikes in the CREDS note "
author: "Ian Philps i.philips@leeds.ac.uk, Jillian Anable J.L.Anable@leeds.ac.uk"
date: "02/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How much e-bike induced mileage would be needed to cancel out the CO2 reduction capablity of e-bikes? 

The mean car km replacement capabilty per person per year is 
~5000

ebikes are almost 8 times more efficient than a petrol car.  
CREDS note box 1:  

Lifecycle CO2 emissions g/km
E-bike	22
Battery electric car (Nissan leaf)	104 
hybrid-car (Toyota Prius) 	168
Petrol car (EU average)	258


if each car km replaced induces 7.63 further km the total induced travel is: 

```{r}
168/22

5000* 7.63 

```


## Compare induced km to a well known journey 
The distance from London to Leeds return is 194 miles 

```{r}
(194*2) * 1.602

```

This is the equivalent of 61 return trips which is more than once per week

```{r}
(5000* 7.63 )/ ((194*2) * 1.602)
```

